\Sof{TBA - New Titles in the comments}

### 2020-04-29 paper/icsme20.tex
results

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}
\usepackage{footnote}
\usepackage{xfrac}
\usepackage{balance}
\usepackage{hyperref}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of $10$ 
guidelines using a dataset including $1335$ security commits
\Sof{These are the initial values, should I use the final? In the end, 
we studied 1031 commits and 8 guidelines}. 
Results show 
that fixing software vulnerabilities may hinder the maintainability of 
open-source software. This is particularly the case for patches that 
address Path Traversal (CWE-22), Information Exposure (CWE-200) and Resource 
Management Errors (CWE-399). We conclude that changes to code bases to patch 
vulnerabilities need to be performed with extra care and that security-related 
code evolution should be subject in computer curricula.
\end{abstract}

@@

\subsection{Statistical Validation}\label{sec:statsval}
%
To validate the maintainability differences in different groups of commits
(e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank
test with the significance level $\alpha = 0.05$~\cite{10.2307/3001968}. In
other words, we test the null hypothesis that the maintainability difference
between pairs of versions $v_{s-1}$, $v_s$ (i.e., before and after a
security-commit) come from the same distribution. Nevertheless, this test 
has a limitation: it does not consider the groups of commits with a 
zero-difference maintainability. In $1959$, Pratt provided an improvement to 
the test to solve this issue making the test more robust. Thus, we use a 
version of the Wilcoxon test that incorporates the cases where 
maintainability is different from zero~\cite{10.2307/2282543}.
The Wilcoxon test requires a distribution size of at leat $20$ instances.
To understand the effect-size, as advocated by the Common-language effect 
sizes\cite{graw:1992}, we compute the mean difference, the median of the 
difference, and the percentage of cases that reduce maintainability.
%
\section{Results}\label{sec:results}

This study evaluates $1031$ security and baseline commits from $265$ 
distinct open-source projects. The following section presents the results 
obtained for each research question.

@@

\begin{figure}[h]
 	\centering
	\vspace{-0.3 cm}
 	\includegraphics[width=0.45\textwidth]{figures/main_per_guideline.pdf}
	\vspace{-0.3 cm}
 	\caption{Impact of the security patches per guideline.}
	\label{fig:guidelines}
	\vspace{0.2 cm}
	\scalebox{0.5}{
	\begin{tabular}{r@{: }l r@{: }l}
	$WCC$ & Write Clean Code & $KACB$ & Keep Architecture Components Balanced\\
	$CACL$ & Couple Architecture Components Loosely & $SCM$ & Separate Concerns in Modules \\
	$KUIS$ & Keep Unit Interfaces Small & $WCO$ & Write Code Once \\
	$WSimpleUC$ & Write Simple Units of Code & $WShortUC$ & Write Short Units of Code
	\end{tabular}
	}	
\end{figure}

@@

Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $6$ guidelines: 
\emph{Write Clean Code} (p-value = $5.49$x$10^{-54}$), \emph{Keep 
Architecture Components Balanced} (p-value = $5.30$x$10^{-71}$), 
\emph{Couple Architecture Components Loosely} (p-value = $2.24$x$10^{-07}$), 
\emph{Separate Concerns in Modules} (p-value = $0.399$), \emph{Keep Unit 
Interfaces Small} (p-value = $1.94$x$10^{-13}$), \emph{Write Code Once} 
(p-value = $1.03$x$10^{-34}$).

However, for each guideline we also observe a 
substantial number of cases where the maintainability suffers a negative 
impact $-$ from $10\%$ to $40\%$ $-$ which highlights the need of better 
practices/techniques/tools to produce software. \emph{Separate Concerns in 
Module} guideline does not the p-value is higher than $0.05$.

According to the 
results, \emph{Write Simple Units of Code} (p-value=$0.015$ and $38.2\%$/$396$ 
of negative cases) and \emph{Write Short Units of Code} (p-value=$0.029$ and 
$38.3\%$/$397$ of negative cases) are the most affected affected guidelines.

 %%%%%%%%%%% HERE

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/main_per_severity.pdf}
 	\caption{Maintainability difference by the vulnerabilities severity.}
	\label{fig:severity}
\end{figure}

@@

For the vulnerabilities identified with CVE(s), we scraped the National Vulnerability Database (NVD) Website to retrieve 
their severity levels. In total, we obtained severity scores for $574$ 
vulnerabilities: $121$ \emph{HIGH}, $423$ \emph{MEDIUM} and $30$ \emph{LOW}. 
Figure~\ref{fig:severity} presents the impact
of security patches per severity level on the maintainability of open-source 
software. Although with no statistical significance, we observe that 
vulnerabilities with higher severity levels (\emph{HIGH} and \emph{MEDIUM}) have 
more negative impact than the \emph{LOW} severity level on the
maintainability of software.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/main_per_language.pdf}
  \caption{Maintainability difference by language after security patches.}
  \label{fig:lang_main}  
\end{figure}

We have also evaluated the impact on the maintainability per programming language
of open-source software, Figure~\ref{fig:lang_main}. We restrict this analysis to 
programming languages with at least $20$ data points, as this is a requirement 
for our hypothesis tests. Thus, we compare the results for \emph{Python}, \emph{C/C++}, \emph{Ruby}, \emph{PHP}, 
\emph{Objective-C/C++}, and \emph{Java}. The other languages were
integrated into \emph{Others}. The results of the Paired Wilcoxon signed-rank
test by language indicate that there is statistical evidence to support that the
maintainability of open-source software decreases for security patches for \emph{C/C++} 
(p-value = $2.5$x$10^{-5}$). 
% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.
\emph{Ruby} and \emph{PHP} languages also impact negatively the maintainability of software. 
Despite not having statistical significance, there are a considerable number of \emph{Ruby} 
and \emph{PHP} patches hindering maintainability that should not be disregarded.

Overall, it is important to notice that there is a very significant number of patches that
hinder the maintainability of software $-$ even if the number of negative cases does not 
surpass the number of positive or none $-$ for  $3$ different languages: \emph{Python},
\emph{Objective-C/C++} and \emph{Java}. For those languages, the negative impact is between $35\%$ and $50\%$ which highlights the need for better and secure programming languages. 
Results should be read with care for the languages where statistical significance was not found. However, 
data reports very interesting hints about the impact of programming languages in 
security patches.

Regarding security commits results, we observe that maintainability decreases in
$41.61\%$ ($429$), remains the same in $19.69\%$ ($203$) and increases in $38.70\%$
($399$) of the cases after security patches are applied. The resulting
p-value of the Paired Wilcoxon signed-rank test is $4.9$x$10^{-2}$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact on the maintainability of
open-source software}.

The impact of each type of security patches on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. Results demonstrate evidence that the maintainability of open-source software decreases in $3$ different patterns: \emph{Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (CWE-22)}, \emph{Information Exposure (CWE-200)} and \emph{Resource Management Errors (CWE-399)}.

We also observe that maintainability is less
negatively impacted when patching \emph{SQL Injection (CWE-89)}, \emph{Improper Restriction of XML External Entity Reference (CWE-611)}, \emph{Improper Input Validation (CWE-20)}, \emph{Cross-Site Request Forgery (CWE-352)}, \emph{Permissions, Privileges, and Access Controls (CWE-264)}. However,
for all the previous patterns there is a percentage of security patches that hinder maintainability between $30\%$ and $50\%$ which is very significant. \emph{Cross-Site Scripting (CWE-79)} is the pattern that suffers less positive/negative impact from security patches, i.e., security
patches of this type of vulnerability typically have no impact on the maintainability of open-source software.

Software with \emph{Resource Management Errors (CWE-399)} flaws are the most affected ones by security patches: $54.5\%$ ($12$) of the patches had a negative impact
in the maintainability while $36.4\%$ ($8$) yielded a positive impact, and only
$9.1\%$ ($2$) did not suffer any change. The results also show that fixing
\emph{Information Exposure (CWE-200)} vulnerabilities affects the maintainability of $51.2\%$ ($20$) of the cases
and leaves only $2.6\%$ ($1$) of the cases unaffected. In $46.2\%$ ($18$) of the
cases, maintainability increased. 
Patching \emph{Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (CWE-22)} vulnerabilities has a negative
impact of $48.6\%$ ($18$) in the maintainability of open-source software while
$46.0\%$ ($17$) of the cases exhibit maintainability improvement and $5.4\%$
($2$) show that maintainability does not suffer any change. The
\emph{Miscellaneous} patches pattern integrates several types of security
flaws. These type of patches harm maintainability in $41.3\%$ ($673$) of the
cases. In $22.7\%$ ($153$), there is no impact and for the
other $36.0\%$ ($242$) maintainability increases.
It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting (CWE-79)} vulnerabilities. \emph{Cross-Site Scripting (CWE-79)}
does not suffer any decrease in the maintainability of $38.7\%$ ($24$) of the cases
and $35.5\%$ ($22$) have a negative effect on it. Only $25.8\%$ ($16$) of the
cases improve the software maintainability.

The results of the
Paired Wilcoxon signed-rank per pattern exhibited statistical
evidence supporting that the maintainability of open-source software decreases for
only one category: \emph{Miscellaneous} (p-value = $0.033$). However, no statistical
significance was found to the other patterns. Thus, results should be read with care.

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.4\textwidth]{figures/main_per_cwe.pdf}
 	\caption{Maintainability difference by pattern after security patches.}
	\label{fig:pat}	
\end{figure}

\subsubsection*{\textit{\textbf{RQ3} \textbf{What is the impact of security patches versus regular changes on the
maintainability of open-source software?}}}

The impact of security and baseline commits in the maintainability of
open-source software is presented in Figure~\ref{fig:secvsreg}.

\begin{figure}[h]	
 	\centering 	\includegraphics[width=0.43\textwidth]{figures/main_comparison.pdf}
 	\caption{Maintainability difference for security and baseline patches.}
	\label{fig:secvsreg}
\end{figure}

We have seen previously a deterioration in maintainability when patching security vulnerabilities. 
For regular commits, we observe that the maintainability decreases in $21.04\%$ ($133$) 
of the cases after regular changes are applied. But in contrast
to security commits, it increases in $26.90\%$ ($170$) 
and remains the same in $52.06\%$ ($329$) 
of the cases. The resulting p-value of the Paired Wilcoxon signed-rank
test is $5.74x10^{-78}$. The baseline is still being computed at the time of submission (it is computationally very expensive, and will be concluded soon) but it is possible to observe 
that regular changes typically have little impact on the maintainability which contradicts research performed in the past. However,
we manually inspected a considerable amount of cases where this happens and
found that identifying regular commits as similar in size as the security-related commit is limiting the type of regular commits (e.g, input refactorings, renaming variables or functions)\footnote{Example of a random regular commit generated with our methodology (Accessed on \today{}):
\url{https://github.com/AFNetworking/AFNetworking/compare/9cde4e45845486bdb46ad322448f93df3b8479bf...3486a008a1fd1348d0f2fe748a5e23933d94647a}}.

%Since the p-value is below the significance level of $0.05$, 
%we argue that \emph{regular refactorings have a
%positive impact in the maintainability of open-source software}.

@@

\subsubsection*{\textit{\textbf{RQ1} \textbf{What is the impact of security patches on the maintainability
of open-source software?}}}

\textit{Security patches harm the maintainability of open-source software.}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}
\usepackage{footnote}
\usepackage{xfrac}
\usepackage{balance}
\usepackage{hyperref}
\usepackage{array}
\usepackage{subfigure}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of $10$ 
guidelines using a dataset including $1335$ security commits.
Results show that fixing software vulnerabilities may hinder the maintainability of 
open-source software. This is particularly the case for patches that 
address Path Traversal (CWE-22), Information Exposure (CWE-200) and Resource 
Management Errors (CWE-399). We conclude that changes to code bases to patch 
vulnerabilities need to be performed with extra care and that security-related 
code evolution should be subject in computer curricula.
\end{abstract}

@@

\subsection{Statistical Validation}\label{sec:statsval}
%
To validate the maintainability differences in different groups of commits
(e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank
test with the significance level $\alpha = 0.05$~\cite{10.2307/3001968}. In
other words, we test the null hypothesis that the maintainability difference
between pairs of versions $v_{s-1}$, $v_s$ (i.e., before and after a
security-commit) come from the same distribution. Nevertheless, this test 
has a limitation: it does not consider the groups of commits with a 
zero-difference maintainability. In $1959$, Pratt provided an improvement to 
the test to solve this issue making the test more robust. Thus, we use a 
version of the Wilcoxon test that incorporates the cases where 
maintainability is different from zero~\cite{10.2307/2282543}.
The Wilcoxon test requires a distribution size of at leat $20$ instances.
To understand the effect-size, as advocated by the Common-language effect 
sizes\cite{graw:1992}, we compute the mean difference, the median of the 
difference, and the percentage of cases that reduce maintainability.
%
\section{Results \& Discussion}\label{sec:results}

This study evaluates a final total of $998$ security and baseline commits from $262$ 
distinct open-source projects. The following section presents the results 
obtained for each research question.

@@

@@

Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $6$ guidelines: 
\emph{Write Clean Code} (p-value = $3.66$x$10^{-51}$), \emph{Keep 
Architecture Components Balanced} (p-value = $5.07$x$10^{-73}$), 
\emph{Couple Architecture Components Loosely} (p-value = $1.61$x$10^{-08}$), 
\emph{Separate Concerns in Modules} (p-value = $0.090$), \emph{Keep Unit 
Interfaces Small} (p-value = $8.76$x$10^{-15}$), \emph{Write Code Once} 
(p-value = $4.97$x$10^{-35}$). Overall, for each of the $8$ guidelines we also 
observe a substantial number of cases where the maintainability suffers a 
negative impact $-$ from $10\%$ to $40\%$ $-$ which highlights the need of 
better practices/techniques/tools to produce software. However, the results do 
not provide statistical significance to the \emph{Separate 
Concerns in Modules} guideline - results should be read carefully.

According to the results, the \emph{Write Simple Units of Code} 
(p-value=$0.037$ and $372$ ($37.3\%$) of negative cases) and \emph{Write Short 
Units of Code} (p-value=$0.028$ and $375$ ($37.6\%$) of negative cases) are 
the most affected guidelines.

@@

 \begin{figure}[h]
  	\centering
 	\vspace{-0.3 cm}
  	\includegraphics[width=0.35\textwidth]{figures/main_per_guideline.pdf}
 	\vspace{-0.3 cm}
  	\caption{Impact of the security patches per guideline.}
 	\label{fig:guidelines}
 	\vspace{0.2 cm}
 	\scalebox{0.5}{
 	\begin{tabular}{r@{: }l r@{: }l}
 	$WCC$ & Write Clean Code & $KACB$ & Keep Architecture Components Balanced\\
 	$CACL$ & Couple Architecture Components Loosely & $SCM$ & Separate Concerns in Modules \\
 	$KUIS$ & Keep Unit Interfaces Small & $WCO$ & Write Code Once \\
 	$WSimpleUC$ & Write Simple Units of Code & $WShortUC$ & Write Short Units of Code
 	\end{tabular}
 	}	
 \end{figure}

Some of the vulnerabilities are identified with \emph{Common Vulnerabilities 
and Exposure} (CVE) entries. We leveraged the \emph{National Vulnerability 
Database} (NVD) website to obtain their severity levels. In total, we 
retrieved severity scores for $551$ vulnerabilities: $117$ (\emph{HIGH}), 
$404$ (\emph{MEDIUM}) and $30$ (\emph{LOW}). Figure~\ref{fig:severity} 
presents the impact of security patches per severity level on the 
maintainability of open-source software. We observe that patches for 
\emph{HIGH} ($47.9\%$) and \emph{MEDIUM} ($44.1\%$) severity vulnerabilities 
hinder more the maintainability of software than \emph{LOW} ($30.0\%$) 
severity vulnerabilities. Again, patches have a considerable 
negative impact on software maintainability - between $30\%$ 
and $50\%$. None statistical significance was retrieved. However, 
results should not be disregarded because they seem to confirm the 
assumption that higher severity vulnerabilities patches may have a 
more negative impact on maintainability.

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.4\textwidth]{figures/main_per_severity.pdf}
 	\caption{Maintainability difference by vulnerability severity.}
	\label{fig:severity}
\end{figure}

We have also evaluated the impact on the maintainability per programming 
language of open-source software (Figure~\ref{fig:lang_main}). We restrict 
this analysis to programming languages with at least $20$ data points, as this 
is a requirement for our hypothesis tests. Thus, we compare the results for \texttt{C/C++}, 
\texttt{Ruby}, \texttt{Java}, \texttt{Objective C/C++}, \texttt{Python} and \texttt{PHP}. 
The remaining languages are included into \texttt{Other}.

\texttt{C/C++} ($46.54\%$), \texttt{Ruby} ($46.48\%$) and \texttt{PHP} ($38.27\%$) are
the programming languages with a more negative impact on the maintainability 
of software. \texttt{Java} seems to be the language less affected
by patching, more than $50\%$ of patches had a positive impact on maintainbility. 
But overall patches in all languages have a considerable
negative impact on maintainability - from $30\%$ to $50\%$ - which confirms
the need for better and more secure programming languages. 
Statistical significance was only retrieved for the \texttt{C/C++} 
(p-value = $2.24$x$10^{-05}$) and \texttt{Ruby} (p-value = $0.041$) languages.
Results should be read with care where statistical significance was not found.
Yet, data reports very interesting hints on the impact of programming languages on 
security patches.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.4\textwidth]{figures/main_per_language.pdf}
  \caption{Maintainability difference by programming language.}
  \label{fig:lang_main}  
\end{figure}

% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.

Regarding security commits results (Figure~\ref{fig:secvsreg}), we observe that
maintainability decreases in $41.28\%$ ($412$), remains the same in $20.64\%$ ($206$)
and increases in $38.08\%$ ($380$) of the cases. The resulting
p-value of the Paired Wilcoxon signed-rank test is $0.044$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact on the maintainability of
open-source software}.
%

\textit{\textbf{RQ2: Which weaknesses are more likely to
affect open-source software maintainability?}}

\begin{figure*}[htp]
  \centering
  \subfigure[Maintainability difference by first-level weaknesses from the 
  \textit{Research Concepts} list on Common Weakness Enumeration (CWE)]{\includegraphics[scale=0.4]{figures/main_per_cwe.pdf}}\quad\quad
  \subfigure[Maintainability difference by sub-weaknesses of the 
  \textit{Improper Neutralization} Weakness (CWE-707)]{\includegraphics[scale=0.4]{figures/main_per_cwe_spec_707.pdf}}\quad\quad
    \subfigure[Maintainability difference by sub-weaknesses of the 
	\textit{Improper Control of a Resource Through its Lifetime} Weakness (CWE-664)]{\includegraphics[scale=0.4]{figures/main_per_cwe_spec_664.pdf}}
	\label{fig:pat}
\end{figure*}

\emph{Weakness}, according to Common Weakness Enumeration (CWE) glossary, 
is a type of mistake that, in proper conditions, could contribute to the 
introduction of vulnerabilities within that product. This enumeration
system was used to classify the vulnerabilities included in our study.
Figure~\ref{fig:pat} shows three different charts: \emph{(a)} shows
the impact of all patches grouped using the first-level weaknesses from
the \emph{Research Concepts} list on CWE; and, the remaining charts
show two of those first-level weaknesses more in depth reporting the impact 
on maintainability of their sub-weaknesses (\emph{(b) and (c)}).

Considering the first-level weaknesses from the \emph{Research Concepts}
list (Figure~\ref{fig:pat}-\emph{a}), we observe that patching vulnerabilities
have negative impact in $5$ weaknesses: Insufficient Control Flow Management 
(CWE-691), Protection Mechanism Failure (CWE-693), Improper Access Control (CWE-284), 
Improper Neutralization (CWE-707) and Improper Control of a Resource Through its Lifetime
(CWE-664). The CWE-707 ($301$) and CWE-664 ($321$) weaknesses integrate a 
larger amount of cases compared to the remaining weaknesses.
Thus, we present an analysis of their sub-weaknesses on Figure~\ref{fig:pat}-\emph{b} and
Figure~\ref{fig:pat}-\emph{c}, respectively. Results demonstrate evidence that patching vulnerabilities hinders 
the maintainability of open-source software for $5$ different weaknesses: \emph{Improper Input Validation 
(CWE-20)}, \emph{Information Exposure (CWE-200)}, \emph{Improper Restriction of Operations within the Bounds of 
a Memory Buffer (CWE-119)}, \emph{Missing Release of Memory after Effective Lifetime 
(CWE-401)} and \emph{Path Traversal (CWE-22)}. Results also show that software maintainability
is less negatively impacted when patching \emph{Improper Restriction of XML External Entity Reference 
(CWE-611)}.

\emph{Cross-Site Scripting (CWE-79)} is the weakness that suffers less 
positive/negative impact from security patches, i.e., patching usually 
has no impact on the maintainability of open-source software. This can 
be explained by the fact that patches to this weakness usually are easily 
done by adding an \texttt{escape} function. Patches of \emph{SQL Injection
(CWE-89)} weaknesses equally hinder and improve the software maintainability. 
But overall a very considerable percentage of patches hinder the open-source 
software maintainability - between $30\%$ and $60\%$.

Statistical significance was only found to the CWE-119 weakness (p-value = 
$0.041$). However, results obtained to the other weaknesses should not be 
disregarded because they offer an interesting view on how the maintainability 
is impacted per weakness which can provide insights to developers/researchers
on the weaknesses that need more attention.

\textit{\textbf{RQ3: What is the impact of security patches versus regular patches on the
maintainability of open-source software?}}

The impact of security and baseline changes in the maintainability of
open-source software is presented in Figure~\ref{fig:secvsreg}.  
We have seen, previously, a deterioration on software maintainability 
when patching vulnerabilities: $41.28\%$ ($412$) suffered a 
negative impact, $20.64\%$ ($206$) remained the same and $38.08\%$ 
($380$) increased the software maintainability. For regular changes, 
we observe that the maintainability decreases in $27.05\%$ ($270$) 
and increases in $30.26\%$ ($302$) of the cases. But in contrast to 
security patches, the maintainability of regular changes remains the 
same in $42.69\%$ ($426$) of the cases and performing regular
changes has a more positive impact than negative on the software 
maintainbility.

Results show that regular changes typically have no impact on the 
software maintainability which contradicts previous research.
However, we manually inspected a considerable amount of cases from
the regular changes distribution where this happens and found that 
identifying regular commits as similar in size as the security-related 
commit is limiting the type of regular commits (e.g, input 
refactorings, variables or functions, type conversion, etc).  
One example is the regular commit\footnote{Regular change: \url{https://github.com/AFNetworking/AFNetworking/compare/9cde4e45845486bdb46ad322448f93df3b8479bf...3486a008a1fd1348d0f2fe748a5e23933d94647a}} generated 
by our methodology and using as reference the following security commit
\footnote{Security change: \url{https://github.com/AFNetworking/AFNetworking/compare/49186d9a3032dc02da1d0cc445cbf69b0cfb001b...2b2558a90e7bfb2ef5169b1ca21b91851e5a5b7d}}. From a
overflow vulnerability patch, our methodology chose a similiar
type of change. We strongly believe that this phenomenon lead
that increased significantly the number of cases with no impact
on the software maintainability.

The resulting p-value of the Paired Wilcoxon signed-rank test 
is $0.331$, i.e., conclusions regarding the regular changes 
are not statistically significant. Yet, it is important to
note that \emph{patching security vulnerabilities seem to have a 
more negative impact on the maintainability of open-source 
software than performing regular changes}.

\begin{figure}[h]	
 	\centering 	\includegraphics[width=0.4\textwidth]{figures/main_comparison.pdf}
 	\caption{Maintainability difference for security and baseline patches.}
	\label{fig:secvsreg}
\end{figure}

@@

\textit{\textbf{RQ1: What is the impact of security patches on the maintainability
of open-source software?}}
\textit{Patching vulnerabilities may harm the maintainability of open-source software.}
